---
title: "Diabetic Retinopathy - Final Project"
author: "Hannah Kim"
date: "8/17/2021"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

## Introduction

This project uses the Diabetic Retinopathy dataset from the UCI Machine Learning Repository. 
Theses are the physician-scientists that collected the datapoints: 

- Dr. Balint Antal, Department of Computer Graphics and Image Processing
Faculty of Informatics, University of Debrecen, 4010, Debrecen, POB 12, Hungary
antal.balint '@' inf.unideb.hu

- Dr. Andras Hajdu, Department of Computer Graphics and Image Processing
Faculty of Informatics, University of Debrecen, 4010, Debrecen, POB 12, Hungary
hajdu.andras '@' inf.unideb.hu

This dataset was collected in 11/03/2014.

## General Dataset Information

Currently, the dataset contains 20 attributes or features that are numbered from X1 to X20.

```{r cars}
library(readr)
messidor_features <- read_csv("Desktop/messidor_features.csv", col_names = FALSE)
colnames(messidor_features)
```

Going from X1 to X20, the attributes or columns represent: 
Image quality assessment
Pre-screening assessment
Microaneurysm detection in 6 different lighting conditions(col 3-8) 
Exudate detection in 8 different lighting conditions (col 9-16)
Distance between macula and optic disc
Optic disc diameter
Class label based on AM/FM
Class label (Diagnosis of diabetic retinopathy)

## Summary of libraries used

Readr (Data importing)
Tidyverse (Data cleaning)
Ggplot2 & pastecs (Assessing the normalization of features)
RColorBrewer (Using colorblind friendly palettes)

```{r cars}
library(tidyverse)
library(ggplot2)
library(pastecs)
library(RColorBrewer)
```

## Data Cleaning

To organize the multiple data points about the number of microaneurysms present in one fundus photo, I used the mutate() function compute the average and place the values at the end of the table.
```{r cars, echo = TRUE}
messidor_features <- messidor_features %>% mutate((X3+X4+X5+X6+X7+X8)/6)
```

Similarly, to organize the multiple data points about the number of exudates present in one fundus photo, I used the mutate() function compute the average and place the values at the end of the table.

```{r cars, echo = TRUE}
messidor_features <- messidor_features %>% mutate((X9+X10+X11+X12+X13+X14+X15+X16)/8)
```

The columns that were used to compute the average was then removed.
```{r cars, echo = TRUE}
DiabRet <- messidor_features[ -c(3:16) ]
```

Then the columns were renamed appropriately to reflect the description of the attribute or column.
```{r cars, echo = TRUE}
DiabRet <- rename(DiabRet, ImageQuality = X1)
DiabRet <- rename(DiabRet, Prescreening = X2)
DiabRet <- rename(DiabRet, DistanceOD_Macula = X17)
DiabRet <- rename(DiabRet, ODDiameter = X18)
DiabRet <- rename(DiabRet, ClassAMFM = X19)
DiabRet <- rename(DiabRet, ClassFinal = X20)
DiabRet <- rename(DiabRet, AvgMicroaneurysms = "(X3 + X4 + X5 + X6 + X7 + X8)/6")
DiabRet <- rename(DiabRet, AvgExudates = "(X9 + X10 + X11 + X12 + X13 + X14 + X15 + X16)/8")
```

## Data Summary

This is a summary of the data:
```{r cars, echo = TRUE}
summary(DiabRet)
```
As you can see there is a greater spread of data with the avg microaneurysms and avg exudates values compared to the distance between OD to macula values or the OD diameter values.

## Data Visualization with Histograms

As taught in class, histograms enable us to visualize the spread of data.
```{r pressure}
hist(DiabRet$DistanceOD_Macula)
```

```{r pressure}
hist(DiabRet$ODDiameter)
```

```{r pressure}
hist(DiabRet$AvgMicroaneurysms)
```

```{r pressure}
hist(DiabRet$AvgExudates)
```

This is the histograms of the 4 main variables or attributes in the dataset. 
The OD diameter variable seems to be the most normally distributed.
Then the distance between OD to macula variable are the next most normally distributed.
Both the avg microaneurysm and avg exudates variables are skewed to the right.

## Data Visualization with Geompoint()

This plot compares the avg microaneurysms (located in the x axis) to the avg exudates (located on the y axis)
The purple dots are images that were diagnosed with diabetic retinopathy.
The orangle red dots are images that were normal/healthy retinas.
```{r pressure}
ggplot(DiabRet, aes(x = AvgMicroaneurysms, y = AvgExudates, col = ClassFinal))+geom_point() + scale_color_brewer(breaks = 3,palette = "Dark2")
```

This plot compares the distance between OD to macula variable (located on the x) to the OD diameter variable (located on the y)
```{r pressure}
ggplot(DiabRet, aes(x = DistanceOD_Macula, y = ODDiameter, col = ClassFinal))+geom_point() + scale_color_brewer(breaks = 3,palette = "Dark2")
```

## Multiple Regression Model

The multiple regression model was used to observe how different variables affect the number of microaneurysms present or detected in the retina. 
```{r pressure}
DRModel <- lm(AvgMicroaneurysms ~ AvgExudates + DistanceOD_Macula + ODDiameter + ClassFinal, data = DiabRet)
summary(DRModel)
```

In this model, the dependent variable is the number of microaneurysms. 
The independent variables are the other 3 main variables and the diagnoses.

Looking at the summary table: 
I don’t think any of the variables had a p value lower than 0.05, so they were not statistically significant enough to make a concrete conclusion.
As seen in the ggplot, the average number of exudates has a similar spread of datapoints to the avg microaneurysms - and here it also shows that it would be the best predictor of microaneurysms when compared to other variables. 
The table reads that an increase of one exudate would contribute to a decrease in about a half of a microaneurysm.

Error Analysis:
This table is tricky to read because an increase of one cm between the distance between OD and the macula would decrease the number of microaneurysms by 23.
This doesn’t make sense because the distance between the OD and the macula will never be greater than 4 to 5 mm.
If all the variables were normalized to have the same scale of values, it would show that this distance variable is largely insignificant. 










